---
title: "Exercises for class 1"
output: html_document
date: "13-02-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
```

# Exercises for Class 1

## Supporting code

Below is code from Chapter 2 of the book that you can use to do the exercises. Not all the code you require is here though. Make sure to not rush it and take time to play around with each function. :))

```{r}
# Grid Approximation

# define grid
p_grid = seq(from = 0, to = 1, length.out=20)

# define prior
prior = rep(1, 20) # each value in the grid has equal probability 

# compute likelihood at each value in grid
likelihood = dbinom(6, size = 9, prob = p_grid) # the likelihood of observing 6 out of 9 trials for each value in the grid, assuming a binomial distribution 

# compute product of likelihood and prior
unstd.posterior = likelihood * prior # unstandardized posterior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")
mtext("20 points")

```

## Exercises

### Easy

2E1. Which of the expressions below correspond to the statement: the probability of rain on Monday?

  (2) Pr(rain∣Monday)

2E2.Which of the following statements corresponds to the expression: Pr(Monday|rain)?

  (3) The probability that it is Monday,given that it is raining.

2E3. Which of the expressions below correspond to the statement: the probability that it is Monday, given that it is raining?

  (1) Pr(Monday|rain)


2E4.

When we say "the probability of water is 0.7", we're not making an assertion about the objective reality of the globe's composition. Instead, we're expressing our subjective degree of belief or confidence based on the available evidence. It's a way to quantify and communicate our uncertainty in light of our current knowledge, rather than making a definitive statement about the true state of the world.

### Medium

2M1.

```{r}

# W, W, W
# define grid
p_grid = seq(from = 0, to = 1, length.out=3) #seq from 0 to 3 as required in WWW

# define prior
prior = rep(1, 3)  

# compute likelihood at each value in grid
likelihood = dbinom(3, size = 3, prob = p_grid) # the likelihood of observing 3 out of 3 trials for each value in the grid, assuming a binomial distribution 

# compute product of likelihood and prior
unstd.posterior = likelihood * prior # unstandardized posterior

# standardize the posterior so it sums to 1
posterior = unstd.posterior / sum(unstd.posterior)

# plot the distribution
plot(p_grid, posterior, type = "b",
     xlab = "probability of water", ylab = "posterior_probability")



```


2M2.
Now assume a prior for p that is equal to zero when p < 0.5 and is a positive constant when p ≥ 0.5. Again compute and plot the grid approximate posterior distribution for each of the sets of observations in the problem just above.

```{r}

# W, W, W, L
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prior <- ifelse(p_grid < 0.5, 0, 1) # 0 when smaller than 0.5, otherwise 1
likelihood <- dbinom(3, 4, prob = p_grid)
unstd.posterior <- likelihood * prior
posterior <- unstd.posterior / sum(unstd.posterior)
plot(p_grid, posterior, type = "l", xlab = "Probability of water", ylab = "Posterior probability", col = "blue")

```

2M3.

Pr(earth|land)= Pr(land|earth)*Pr(earth)/Pr(land)
                   0.3             0.5         0.65
```{r}
likelihood <- 0.3
prior <- 0.5
normalising_constant <- 0.65

posterior <- likelihood*prior/normalising_constant


```

For the exercises below, I highly suggest you to grab a piece of paper and try to solve the problems in a 'visual' way.

2M4. Suppose you have a deck with only three cards. Each card has two sides, and each side is either black or white. One card has two black sides. The second card has one black and one white side. The third card has two white sides. Now suppose all three cards are placed in a bag and shuffled. Someone reaches into the bag and pulls out a card and places it flat on a table. A black side is shown facing up, but you don’t know the color of the side facing down. Show that the probability that the other side is also black is 2/3. Use the counting method (Section 2 of the chapter) to approach this problem. This means counting up the ways that each card could produce the observed data (a black side facing up on the table).

```{r}

prob <- 2/3

```

2M5.Now suppose there are four cards: B/B, B/W, W/W, and another B/B. Again suppose a card is drawn from the bag and a black side appears face up. Again calculate the probability that the other side is black.

```{r}

prob <- 4/5

```

2M6.Imagine that the black ink is heavy, and so cards with black sides are heavier than cards with white sides. As a result, it’s less likely that a card with black sides is pulled from the bag. So again assume there are three cards: B/B, B/W, and W/W. After experimenting a number of times, you conclude that for every way to pull the B/B card from the bag, there are 2 ways to pull the B/W card and 3 ways to pull the W/W card. Again suppose that a card is pulled and a black side appears face up. Show that the probability the other side is black is now 0.5. Use the counting method, as before.

```{r}

prob <- 1/2

```
